---
title: "French Media 2013-2020"
subtitle: "4. Networks"
author: "Claude Grasland"
output: html_notebook
---


```{r setup4, echo = FALSE, comment = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = FALSE, warning = FALSE)
library(knitr)
library(dplyr)
library(data.table)
library(visNetwork, quietly = T)
```




# Networks


## Preparation of data

### Load Hypercube

```{r}
# Load complete hypercube
hc <- readRDS("hypercubes/hc_fr_FRA_geo_geo.RDS")
hc <- hc %>% filter(where1 !="_no_", where2 != "_no_")

# Add french labels
labs<-read.table("dict/worldgeo_label_fr.csv", sep=";",header=T, encoding = "UTF-8")
labs<-labs[,c(5,3)]
names(labs)<-c("where1","geofr1")
hc<-left_join(hc,labs)
names(labs)<-c("where2","geofr2")
hc<-left_join(hc,labs)

hc_geo_geo <- hc
```


### Hypercube Filter (function)

```{r}
hc_filter <- function(don = hc,
                      who = "who",
                      when = "when",
                      where1 = "where1",
                      where2 = "where2",
                      wgt = "tags",
                      self = FALSE,
                      when_start = NA,
                      when_end = NA,
                      who_exc = NA,
                      who_inc = NA,
                      where1_exc = NA,
                      where1_inc = NA,
                      where2_exc = NA,
                      where2_inc = NA)

  {                          
  
    df<-data.table(who = don[[who]],
                   when = don[[when]],
                   where1 = don[[where1]],
                   where2 = don[[where2]],
                   wgt = don[[wgt]])
    
    # Select time period
        if (is.na(when_start)==FALSE) { 
        df <- df[when >= as.Date(when_start), ]}
        if (is.na(when_end)==FALSE) { 
        df <- df[when <= as.Date(when_end), ]}
    # Select who
        if (is.na(who_exc)==FALSE) { 
        df <- df[!(who %in% who_exc), ]}
        if (is.na(who_inc)==FALSE) { 
        df <- df[(who %in% who_inc), ]}
    # Select where1
        if (is.na(where1_exc)==FALSE) { 
        df <- df[!(where1 %in% where1_exc), ]}
        if (is.na(where1_inc)==FALSE) { 
        df <- df[(where1 %in% where1_inc), ]}
    # Select where2
        if (is.na(where2_exc)==FALSE) { 
        df <- df[!(where2 %in% where2_exc), ]}
        if (is.na(where2_inc)==FALSE) { 
        df <- df[(where2 %in% where2_inc), ]}
    # eliminate internal links
       if (self==FALSE) { 
        df <- df[(where1 != where2), ]}
    return(df)
  
}
```

### Matrix builder (function)

```{r}
build_int <- function(don = don,       # a dataframe with columns i, j , Fij
                      i = "where1",
                      j = "where2",
                      Fij = "wgt",
                      s1 = 1,
                      s2 = 1,
                      n1 = 1,
                      n2 = 1,
                      k = 0)

{  
  df<-data.table(i=don[[i]],j=don[[j]],Fij=don[[Fij]])
  int <-df[,.(Fij=sum(Fij)),.(i,j)]
  int<-dcast(int,formula = i~j,fill = 0)
  mat<-as.matrix(int[,-1])
  row.names(mat)<-int$i
  mat<-mat[apply(mat,1,sum)>=s1,apply(mat,2,sum)>=s2 ]
  m0<-mat
  m0[m0<k]<-0
  m0[m0>=k]<-1
  mat<-mat[apply(m0,1,sum)>=n1,apply(m0,2,sum)>=n2 ]
  int<-reshape2::melt(mat)
  names(int) <-c("i","j","Fij")
  return(int)
}

```


### Random model (function)

```{r}

rand_int <- function(int = int, # A table with columns i, j Fij
                     maxsize = 100000,
                     diag    = FALSE,
                     resid   = FALSE) {
    # Eliminate diagonal ?
    if (diag==FALSE) { 
        int <- int[as.character(int$i) != as.character(int$j), ]}
  
    # Compute model if size not too large
    if (dim(int)[1] < maxsize) {
       # Proceed to poisson regression model
       mod <- glm( formula = Fij ~ i + j,family = "poisson", data = int)
  
       # Add residuals if requested
       if(resid == TRUE)   { 
          # Add estimates
          int$Eij <- mod$fitted.values

          # Add absolute residuals
          int$Rabs_ij <- int$Fij-int$Eij

          # Add relative residuals
          int$Rrel_ij <- int$Fij/int$Eij

          # Add chi-square residuals
          int$Rchi_ij <-  (int$Rabs_ij)**2 / int$Eij
          int$Rchi_ij[int$Rabs_ij<0]<- -int$Rchi_ij[int$Rabs_ij<0]
          }
         
    } else { paste ("Table > 100000 -  \n 
                     modify maxsize =  parameter \n
                     if you are sure that your computer can do it !")}
  # Export results
  int$i<-as.character(int$i)
  int$j<-as.character(int$j)
  return(int)
  
 }

```


### Visualize network (function)

This standard procedure has to be adapted to particular cases

```{r}
geo_network<- function(don = don,
                       from = "i",
                        to = "j", 
                        size = "Fij",
                        minsize = 1,
                        maxsize = NA,
                        test = "Fij",
                        mintest = 1,
                        loops  = FALSE, 
                        title = "Network")

{
int<-data.frame(i = as.character(don[,from]),
                j = as.character(don[,to]),
                size = don[,size],
                test = don[,test]
                )
if (is.na(minsize)==FALSE) {int =int[int$size >= minsize,]} 
if (is.na(maxsize)==FALSE) {int =int[int$size <= maxsize,]} 
if (is.na(mintest)==FALSE) {int =int[int$test >= mintest,]}

nodes<-data.frame(code = unique(c(int$i,int$j)))
nodes$code<-as.character(nodes$code)
nodes$id<-1:length(nodes$code)
nodes$label<-nodes$code
nodes$color <-"gray"
nodes$color[nodes$code %in% int$j]<-"red"


# Adjust edge codes
edges <- int %>% mutate(width = 5+5*size / max(size)) %>%
                left_join(nodes %>% select(i=code, from = id)) %>%  
                left_join(nodes %>% select(j=code, to = id )) 

# compute nodesize
toti<-int %>% group_by(i) %>% summarize(size =sum(size)) %>% select (code=i,size)
totj<-int %>% group_by(j) %>% summarize(size =sum(size)) %>% select (code=j,size)
tot<-rbind(toti,totj)
tot<-unique(tot)
tot$code<-as.factor(tot$code)
nodes <- left_join(nodes,tot) %>% mutate(value = 1 +5*sqrt(size/max(size)))


#sel_nodes <-nodes %>% filter(code %in% unique(c(sel_edges$i,sel_edges$j)))

# eliminate loops

if(loops == FALSE) {edges <- edges[edges$from < edges$to,]}

net<- visNetwork(nodes, 
                  edges, 
                  main = title,
height = "1000px", 
                  width = "70%")   %>%   
   visNodes(scaling =list(min =20, max=60, 
                          label=list(min=20,max=80, 
                                    maxVisible = 20)))%>%
  visEdges(scaling = list(min=20,max=60))%>%
       visOptions(highlightNearest = TRUE,
     #               selectedBy = "group", 
    #               manipulation = TRUE,
                  nodesIdSelection = TRUE) %>%
        visInteraction(navigationButtons = TRUE) %>%
         visLegend() %>%
      visIgraphLayout(layout ="layout.fruchterman.reingold",smooth = TRUE)

net
 return(net)
 } 

```


## Analyse networks


### Regions x Regions / Le Monde
```{r}
hc<-hc_geo_geo %>% filter(substr(where1,1,3) %in% c("CO_","SE_","LA_","OR_")) %>%
                            filter(substr(where2,1,3) %in% c("CO_","SE_","LA_","OR_")) %>% 
                            filter(who == "FRA_lmonde")



hc<-hc_filter(don = hc,
                             wgt = "tags",
                             where1 = "geofr1",
                             where2 = "geofr2",
                             where1_exc = c("_no_"),
                             where2_exc = c("_no_"),
                             self = FALSE
                           )

int <- build_int(don = hc,
                 s1=2,
                 s2=2,
                 n1=2,
                 n2=2,
                 k=0)

mod<-rand_int(int,
              resid = TRUE,
              diag = FALSE)

sum(mod$Fij)
network<- geo_network(mod,
                      size = "Fij",
                      minsize = sum(mod$Fij/1000),
                      test = "Rchi_ij",
                      mintest = 3.84)
network


visSave(network, "regions_regions_lmonde.html")
```



### regions x Regions / Le Figaro
```{r}
hc<-hc_geo_geo %>% filter(substr(where1,1,3) %in% c("CO_","SE_","LA_","OR_")) %>%
                            filter(substr(where2,1,3) %in% c("CO_","SE_","LA_","OR_")) %>% 
                            filter(who == "FRA_figaro")



hc<-hc_filter(don = hc,
                             wgt = "tags",
                             where1 = "geofr1",
                             where2 = "geofr2",
                             where1_exc = c("_no_"),
                             where2_exc = c("_no_"),
                             self = FALSE
                           )

int <- build_int(don = hc,
                 s1=2,
                 s2=2,
                 n1=2,
                 n2=2,
                 k=0)

mod<-rand_int(int,
              resid = TRUE,
              diag = FALSE)

sum(mod$Fij)
network<- geo_network(mod,
                      size = "Fij",
                      minsize = sum(mod$Fij/1000),
                      test = "Rchi_ij",
                      mintest = 3.84)
network


visSave(network, "regions_regions_figaro.html")
```

### regions x Regions / La Croix
```{r}
hc<-hc_geo_geo %>% filter(substr(where1,1,3) %in% c("CO_","SE_","LA_","OR_")) %>%
                            filter(substr(where2,1,3) %in% c("CO_","SE_","LA_","OR_")) %>% 
                            filter(who == "FRA_figaro")



hc<-hc_filter(don = hc,
                             wgt = "tags",
                             where1 = "geofr1",
                             where2 = "geofr2",
                             where1_exc = c("_no_"),
                             where2_exc = c("_no_"),
                             self = FALSE
                           )

int <- build_int(don = hc,
                 s1=2,
                 s2=2,
                 n1=2,
                 n2=2,
                 k=0)

mod<-rand_int(int,
              resid = TRUE,
              diag = FALSE)

sum(mod$Fij)
network<- geo_network(mod,
                      size = "Fij",
                      minsize = sum(mod$Fij/1000),
                      test = "Rchi_ij",
                      mintest = 3.84)
network


visSave(network, "regions_regions_lacroi.html")
```

### regions x Regions / Libération
```{r}
hc<-hc_geo_geo %>% filter(substr(where1,1,3) %in% c("CO_","SE_","LA_","OR_")) %>%
                            filter(substr(where2,1,3) %in% c("CO_","SE_","LA_","OR_")) %>% 
                            filter(who == "FRA_libera")



hc<-hc_filter(don = hc,
                             wgt = "tags",
                             where1 = "geofr1",
                             where2 = "geofr2",
                             where1_exc = c("_no_"),
                             where2_exc = c("_no_"),
                             self = FALSE
                           )

int <- build_int(don = hc,
                 s1=2,
                 s2=2,
                 n1=2,
                 n2=2,
                 k=0)

mod<-rand_int(int,
              resid = TRUE,
              diag = FALSE)

sum(mod$Fij)
network<- geo_network(mod,
                      size = "Fij",
                      minsize = sum(mod$Fij/1000),
                      test = "Rchi_ij",
                      mintest = 3.84)
network


visSave(network, "regions_regions_libera.html")
```


### States x Regions / 4 medias
```{r}
hc<-hc_geo_geo %>% filter(!(substr(where1,1,3) %in% c("CO_","SE_","LA_","OR_"))) %>%
                            filter(substr(where2,1,3) %in% c("CO_","SE_","LA_","OR_")) 




hc<-hc_filter(don = hc,
                             wgt = "tags",
                             where1 = "geofr1",
                             where2 = "geofr2",
                             where1_exc = c("_no_", "France"),
                             where2_exc = c("_no_"),
                             self = FALSE
                           )

int <- build_int(don = hc,
                 s1=2,
                 s2=2,
                 n1=2,
                 n2=2,
                 k=0)

mod<-rand_int(int,
              resid = TRUE,
              diag = FALSE)

sum(mod$Fij)
network<- geo_network(mod,
                      size = "Fij",
                      minsize = sum(mod$Fij/1000),
                      test = "Rchi_ij",
                      mintest = 3.84,
                      loops = TRUE)
network


visSave(network, "states_regions_FRA.html")
```





